---
title: "Analysis_features"
author: "Laura GREGOIRE"
date: "2023-11-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r package and path,  include=FALSE}
###packages
library("Publish")
library("car")
library("multcomp")
library("ggplot2")
library("lubridate")
library("stringr")
library("formattable")
library("ggpubr")
library("dplyr")
library("reshape2")
library("tidyr")

getwd()
setwd(dir = "C:/Users/2021lg003/Documents/LEASYSCAN_EXP_INDE_2023/RESULTS/3-AFTER_SMOOTHING/")
path="C:/Users/2021lg003/Documents/LEASYSCAN_EXP_INDE_2023/RESULTS/3-AFTER_SMOOTHING/"

```

```{r load TRrate data }
inputfile_TR<-read.csv("C:/Users/2021lg003/Documents/LEASYSCAN_ANALYSIS/2-DATA_SMOOTHING/results/OP-10_TR_smth.csv", sep = ";", header = TRUE)

colnames(inputfile_TR)[10:ncol(inputfile_TR)] <- sub("X","",colnames(inputfile_TR)[10:ncol(inputfile_TR)])
inputfile_TR=inputfile_TR[-c(8:9)]
timeline=colnames(inputfile_TR[8:ncol(inputfile_TR)])
timeline=dmy_hm(timeline)

##Liste genotypes -trt 
inputfile_TR$Geno_Trt<-str_c(inputfile_TR$Genotype, "_", inputfile_TR$Treatment) ##concatenate Geno et rep
geno_info<-data.frame(inputfile_TR$unit, inputfile_TR$Treatment, inputfile_TR$Genotype, inputfile_TR$Replicates, inputfile_TR$Geno_Trt)
colnames(geno_info)<-c("unit", "Treatment", "Genotype", "Replicates", "Geno_Trt")
geno_info=geno_info[-c(1:8),]
geno_info=na.omit(geno_info)

```

```{r VPD ETref plot}

climate_df<-inputfile_TR[-c(9:nrow(inputfile_TR)),] ###keep only climate rows (removed LC rows)
climate_df<-climate_df[,-c(2:7)] ###enlever les colonnes de 2 a 7
climate_df<-as.data.frame(t(climate_df))
colnames(climate_df)<-climate_df[1,]
climate_df<-climate_df[-1,]
rownames(climate_df)<- sub("X","",rownames(climate_df))
climate_df$Timestamp<-rownames(climate_df)
climate_df <- climate_df %>% mutate_at(c('Temp', 'RH','VPD', 'SR','WS', 'Tmax','Tmin', 'ETref'), as.numeric)
climate_df$Timestamp<-dmy_hm(climate_df$Timestamp)  
climate_df=na.omit(climate_df)

start=which(rownames(climate_df) == "28.09.2023.00.00")
end=which(rownames(climate_df) == "16.10.2023.23.45")
climate_df=climate_df[start:end,] 

mean(climate_df$ETref)
max(climate_df$ETref)
min(climate_df$ETref)
sd(climate_df$ETref)

mean(climate_df$VPD)
max(climate_df$VPD)
min(climate_df$VPD)
sd(climate_df$VPD)

##

  png(file = "VPD_plot.png")
  ggplot(climate_df, aes(x=Timestamp, y=VPD))+
    geom_point() +
    labs(title = "VPD measured") + ylab("VPD (kPa)") + xlab("Date") +
    ggplot2::scale_x_datetime(labels = scales::date_format(format = "%d-%m"), date_breaks = "1 days")+
    theme(axis.text.x=element_text(angle = -45, hjust = 0))+
    geom_line()
  dev.off()
  while (!is.null(dev.list())) 
  
    png(file = "ETref_plot.png")
  ggplot(climate_df, aes(x=Timestamp, y=ETref))+
    geom_point() +
    labs(title = "ETref measured ") + ylab("ETref ((mm 15.min−1)") + xlab("Date") +
    ggplot2::scale_x_datetime(labels = scales::date_format(format = "%d-%m"), date_breaks = "1 days")+
    theme(axis.text.x=element_text(angle = -45, hjust = 0))+
    geom_line()
  dev.off()
  
# max(climate_df$ETref)
#  0.273 mm.15
  
# max(climate_df$VPD)
# 4.970699 kPa
  
# mean(climate_df$ETref)
# 0.05745891
  
# mean(climate_df$VPD)
# 1.462294  

```

```{r TRrate TRIAL PERIOD AND REMOVE IRRGATION DATES}
######### TR rate
TR_profiles<-as.data.frame(t(inputfile_TR))
TR_profiles<-TR_profiles[-c(1:8)] ### remove climate data (first column)
TR_profiles<-TR_profiles[-c(1:2),] 
colnames(TR_profiles)=TR_profiles[nrow(TR_profiles),]
TR_profiles=TR_profiles[-nrow(TR_profiles),]
TR_profiles<-TR_profiles[-c(1:3),]
TR_profiles<-as.data.frame(TR_profiles)
TR_profiles<-TR_profiles[-c(1:2),]
TR_profiles$Timestamp<-rownames(TR_profiles)
TR_profiles$Timestamp<-dmy_hm(TR_profiles$Timestamp)
start=which(rownames(TR_profiles) == "28.09.2023.00.00")
end=which(rownames(TR_profiles) == "16.10.2023.23.45") ## test §§
TR_profiles=TR_profiles[start:end,]

##identify duplicated column : 
geno_info$Geno_Trt[duplicated(geno_info$Geno_Trt)] ## Empty_WW_R4 and NA
geno_info[!duplicated(geno_info$Geno_Trt), ]

## remove duplicated column
TR_profiles <- TR_profiles[, !duplicated(colnames(TR_profiles))]
TR_rate_trial_period=TR_profiles


###removed irrigation dates 

start_irr_day=which(rownames(TR_rate_trial_period) == "30.09.2023.00.00")
end_irr_day=which(rownames(TR_rate_trial_period) == "30.09.2023.23.45")
TR_rate_trial_period=TR_rate_trial_period[-c(start_irr_day:end_irr_day),]

start_irr_day=which(rownames(TR_rate_trial_period) == "01.10.2023.00.00")
end_irr_day=which(rownames(TR_rate_trial_period) == "01.10.2023.23.45")
TR_rate_trial_period=TR_rate_trial_period[-c(start_irr_day:end_irr_day),]

start_irr_day=which(rownames(TR_rate_trial_period) == "03.10.2023.00.00")
end_irr_day=which(rownames(TR_rate_trial_period) == "03.10.2023.23.45")
TR_rate_trial_period=TR_rate_trial_period[-c(start_irr_day:end_irr_day),]

start_irr_day=which(rownames(TR_rate_trial_period) == "06.10.2023.00.00")
end_irr_day=which(rownames(TR_rate_trial_period) == "06.10.2023.23.45")
TR_rate_trial_period=TR_rate_trial_period[-c(start_irr_day:end_irr_day),]

# 
#  start_irr_day=which(rownames(TR_rate_trial_period) == "10.10.2023.00.00")
#  end_irr_day=which(rownames(TR_rate_trial_period) == "10.10.2023.23.45")
#  TR_rate_trial_period=TR_rate_trial_period[-c(start_irr_day:end_irr_day),]
# 
start_irr_day=which(rownames(TR_rate_trial_period) == "13.10.2023.00.00")
end_irr_day=which(rownames(TR_rate_trial_period) == "13.10.2023.23.45")
TR_rate_trial_period=TR_rate_trial_period[-c(start_irr_day:end_irr_day),]


```

```{r TRrate TRIAL PERIOD}
## test i= 1
##TR profil by plot 
for ( i in 1:ncol(TR_rate_trial_period)) { 
  subset_geno_trt <- data.frame(TR_rate_trial_period[i], rownames(TR_rate_trial_period),colnames(TR_rate_trial_period[i]))
  colnames(subset_geno_trt)=c("TRrate", "Timestamp","Geno_Trt")
  subset_geno_trt$Timestamp<-dmy_hm(subset_geno_trt$Timestamp)
  subset_geno_trt$TRrate = as.numeric(subset_geno_trt$TRrate)

  for ( j in 1:nrow(subset_geno_trt)) { 
    
    if(isTRUE(subset_geno_trt$TRrate[j]<0) == TRUE) {
      subset_geno_trt$TRrate[j]=NA
    }
    j=j+1
    print(j)
    }
  
  
  png(file = paste(path,"Transpiration rate Geno_Trt",subset_geno_trt$Geno_Trt[1],".png"))
  
  plot<-
    ggplot(data = subset_geno_trt, aes(subset_geno_trt$Timestamp,subset_geno_trt$TR))   +
    ylab("TRrate (mm m−2 15 min−1)")+
    xlab("Date ")+
    labs(title = subset_geno_trt$Geno_Trt[1])  +
    geom_point()+
    geom_line()   +
    ggplot2::scale_x_datetime(labels = scales::date_format(format = "%d-%m"), date_breaks = "1 days")+
    theme(axis.text.x=element_text(angle = -45, hjust = 0))
  
  print(plot)
  dev.off()
  print(i)

}

```

########### The TR rate profiles are strange: around 0 until October 10, then very high after that. 
Possible cause: TRrate = TR / LA. However, LA should be in NA at the end of handling if the plants have grown a lot and hidden the barcode. (See NA rate in LA dataset per day: over 25% from October 10). So, if LA=NA, instead of taking TRrate, we represent TR at the end and not TRrate at the end of the manipulation. 

To test the hypothesis, we take the OP9 dataset, which is smooth transpiration before LA. Do we have correct profiles?


```{r load TR data }
TR_smooth<-read.csv("C:/Users/2021lg003/Documents/LEASYSCAN_ANALYSIS/2-DATA_SMOOTHING/results/OP-9_ETr_smth.csv", sep = ";", header = TRUE, dec = ".")
colnames(TR_smooth)[9:ncol(TR_smooth)] <- sub("X","",colnames(TR_smooth)[9:ncol(TR_smooth)])
TR_smooth=TR_smooth[-c(1:8),]

## Timestamp
timeline=colnames(TR_smooth[8:ncol(TR_smooth)])
timeline=dmy_hm(timeline)
timeline=na.omit(timeline)

##Liste genotypes -trt 
TR_smooth$Geno_Trt<-str_c(TR_smooth$Genotype, "_", TR_smooth$Treatment) ##concatenate Geno et rep
TR_smooth$Genotype = as.factor(TR_smooth$Genotype )
geno_info<-data.frame(TR_smooth$unit, TR_smooth$Treatment, TR_smooth$Genotype, TR_smooth$Replicates, TR_smooth$Geno_Trt)
colnames(geno_info)<-c("unit", "Treatment", "Genotype", "Replicates", "Geno_Trt")

## change all TR value in character to numeroic 
char_columns <- sapply(TR_smooth, is.character)             # Identify character columns
data_chars_as_num <- TR_smooth                              # Replicate data
data_chars_as_num[ , char_columns] <- as.data.frame(   # Recode characters as numeric
  apply(data_chars_as_num[ , char_columns], 2, as.numeric))
sapply(data_chars_as_num, class)  
TR_trial_period=data_chars_as_num


```

```{r DATASET TRIAL PERIOD AND REMOVE IRRGATION DATES}
######### TR 
TR_profiles<-as.data.frame(t(TR_smooth))
colnames(TR_profiles)=TR_profiles[7,] ##geno_trt as colnames
TR_profiles<-TR_profiles[-c(1:8),] ### remove climate data (first column)



## change all TR value in character to numeroic 
char_columns <- sapply(TR_profiles, is.character)             # Identify character columns
data_chars_as_num <- TR_profiles                              # Replicate data
data_chars_as_num[ , char_columns] <- as.data.frame(   # Recode characters as numeric
  apply(data_chars_as_num[ , char_columns], 2, as.numeric))
sapply(data_chars_as_num, class)  
TR_trial_period=data_chars_as_num



TR_profiles$Timestamp<-rownames(TR_profiles)
TR_profiles$Timestamp <- sub("X","",TR_profiles$Timestamp)
TR_profiles$Timestamp<-dmy_hm(TR_profiles$Timestamp)
TR_profiles$Timestamp=na.omit(TR_profiles$Timestamp)


rownames(TR_profiles) <- sub("X","",rownames(TR_profiles))



start=which(rownames(TR_trial_period) == "28.09.2023.00.00")
end=which(rownames(TR_trial_period) == "16.10.2023.23.45") ## test §§
TR_trial_period=TR_trial_period[start:end,]
# 

###removed irrigation dates 

start_irr_day=which(rownames(TR_trial_period) == "30.09.2023.00.00")
end_irr_day=which(rownames(TR_trial_period) == "30.09.2023.23.45")
TR_trial_period=TR_trial_period[-c(start_irr_day:end_irr_day),]


start_irr_day=which(rownames(TR_trial_period) == "03.10.2023.00.00")
end_irr_day=which(rownames(TR_trial_period) == "03.10.2023.23.45")
TR_trial_period=TR_trial_period[-c(start_irr_day:end_irr_day),]

start_irr_day=which(rownames(TR_trial_period) == "06.10.2023.00.00")
end_irr_day=which(rownames(TR_trial_period) == "06.10.2023.23.45")
TR_trial_period=TR_trial_period[-c(start_irr_day:end_irr_day),]


 # start_irr_day=which(rownames(TR_trial_period) == "10.10.2023.00.00")
 # end_irr_day=which(rownames(TR_trial_period) == "10.10.2023.23.45")
 # TR_trial_period=TR_trial_period[-c(start_irr_day:end_irr_day),]

start_irr_day=which(rownames(TR_trial_period) == "13.10.2023.00.00")
end_irr_day=which(rownames(TR_trial_period) == "13.10.2023.23.45")
TR_trial_period=TR_trial_period[-c(start_irr_day:end_irr_day),]


```

```{r TR per geno_trt TRIAL PERIOD}
## test i= 1
##TR profil by plot 
for ( i in 1:ncol(TR_trial_period)) { 
  subset_geno_trt <- data.frame(TR_trial_period[i], rownames(TR_trial_period),colnames(TR_trial_period[i]))
  colnames(subset_geno_trt)=c("TR", "Timestamp","Geno_Trt")
  subset_geno_trt$Timestamp<-dmy_hm(subset_geno_trt$Timestamp)
  subset_geno_trt$TR = as.numeric(subset_geno_trt$TR)

  for ( j in 1:nrow(subset_geno_trt)) {

    if(isTRUE(subset_geno_trt$TR[j]<0) | (isTRUE(subset_geno_trt$TR[j]> 500 ) ) ) {
      subset_geno_trt$TR[j]=NA
    }
    j=j+1
    print(j)
    }

  
  
  png(file = paste(path,"TR Geno_Trt",subset_geno_trt$Geno_Trt[1],".png"))
  
  plot<-
    ggplot(data = subset_geno_trt, aes(subset_geno_trt$Timestamp,subset_geno_trt$TR))   +
    ylab("TR (mm. 15 min−1)")+
    xlab("Date ")+
    labs(title = subset_geno_trt$Geno_Trt[1])  +
    geom_point()+
    geom_line()   +
    ggplot2::scale_x_datetime(labels = scales::date_format(format = "%d-%m"), date_breaks = "1 days")+
    theme(axis.text.x=element_text(angle = -45, hjust = 0))
  
  print(plot)
  dev.off()
  print(i)

}



```

```{r TR per genotype plot}
##test 
## i=1 ##nombre du level factor 1234
## g = "Exp60_1_IS13"

##before plot, make sure you have removed duplicated column/empty pot


g=geno_info$Genotype
g=as.factor(g)

 for ( i in  1:nlevels(g)) { 
   
   genotype= g[i]
   genotype= as.character(genotype)

    subset_geno=TR_trial_period[,c(colnames(TR_trial_period)[grep(levels(g)[i],colnames(TR_trial_period))])]
    subset_geno= as.data.frame(subset_geno)
    
    if(ncol(subset_geno) > 1) {   ## if we have more than 1 rep 
      
          subset_geno=cbind(subset_geno, rownames(subset_geno))
          subset_geno = melt(subset_geno ,  id.vars = "rownames(subset_geno)", variable.name = 'Geno_Trt')
          colnames(subset_geno)  = c("Timestamp", "Geno_Trt", "TR")
          subset_geno$Timestamp= dmy_hm(subset_geno$Timestamp)
           

             for ( j in 1:nrow(subset_geno)) { 
                if(isTRUE(subset_geno$TR[j]<0 || subset_geno$TR[j]> 1000 ) == TRUE) {
                subset_geno$TR[j]=NA
                }
              j=j+1
              }
          x= subset_geno$Timestamp
          y= subset_geno$TR

           
          plot<- 
          ggplot(subset_geno, aes(x, y)) +
          geom_line(aes(colour = Geno_Trt))+
          ylab("TR (g. 15 min−1)")+
          xlab("Date")+
          labs(title = paste("Transpiration Geno",genotype))  +
          geom_point(aes(color = Geno_Trt))+
          ggplot2::scale_x_datetime(labels = scales::date_format(format = "%d-%m"), date_breaks = "1 days")+
          theme(axis.text.x=element_text(angle = -45, hjust = 0))
    
    
          png(file = paste(path, "Transpiration Geno", genotype,".png"))
          print(plot)
          dev.off()
          
    }
    
    
    
    else  { 
        print("only one repetition")
      
        subset_geno_trt <- data.frame(TR_trial_period[i], rownames(TR_trial_period),colnames(TR_trial_period[i]))
        colnames(subset_geno_trt)=c("TR", "Timestamp","Geno_Trt")
        subset_geno_trt$Timestamp<-dmy_hm(subset_geno_trt$Timestamp)
        subset_geno_trt$TR = as.numeric(subset_geno_trt$TR)

            for ( j in 1:nrow(subset_geno_trt)) {
                if(isTRUE(subset_geno_trt$TR[j]<0) | (isTRUE(subset_geno_trt$TR[j]> 500 ) ) ) {
                subset_geno_trt$TR[j]=NA
                }
            j=j+1
            }

  
  
        png(file = paste(path,"TR Geno_Trt",subset_geno_trt$Geno_Trt[1],".png"))
  
        plot<-
        ggplot(data = subset_geno_trt, aes(subset_geno_trt$Timestamp,subset_geno_trt$TR))   +
        ylab("TR (g. 15 min−1)")+
        xlab("Date ")+
        labs(title = subset_geno_trt$Geno_Trt[1])  +
        geom_point()+
        geom_line()   +
        ggplot2::scale_x_datetime(labels = scales::date_format(format = "%d-%m"), date_breaks = "1 days")+
        theme(axis.text.x=element_text(angle = -45, hjust = 0))
      
        print(plot)
        dev.off()
    }
    
    print(i)
    i=i+1

 }


```



```{r Load features smoothed data }
##### Load features smoothed data on trial period : 
##DO NOT LOAD IRRIGATION DATES
features_d1<-read.csv("C:/Users/2021lg003/Documents/LEASYSCAN_ANALYSIS/2-DATA_SMOOTHING/results/PERIODE1/smthTR_FeaturesTimeSeries/D.2023-09-28.csv", sep = ";")
features_d2<-read.csv("C:/Users/2021lg003/Documents/LEASYSCAN_ANALYSIS/2-DATA_SMOOTHING/results/PERIODE1/smthTR_FeaturesTimeSeries/D.2023-09-29.csv", sep = ";")
#features_d3<-read.csv("C:/Users/2021lg003/Documents/LEASYSCAN_ANALYSIS/2-DATA_SMOOTHING/results/PERIODE2/smthTR_FeaturesTimeSeries/D.2023-09-30.csv") 
#features_d4<-read.csv("./2-DATA_SMOOTHING/results/PERIODE2/smthTR_FeaturesTimeSeries/D.2023-10-01.csv") #
features_d5<-read.csv("C:/Users/2021lg003/Documents/LEASYSCAN_ANALYSIS/2-DATA_SMOOTHING/results/PERIODE2/smthTR_FeaturesTimeSeries/D.2023-10-02.csv") 
features_d6<-read.csv("C:/Users/2021lg003/Documents/LEASYSCAN_ANALYSIS/2-DATA_SMOOTHING/results/PERIODE2/smthTR_FeaturesTimeSeries/D.2023-10-03.csv") 
features_d7<-read.csv("C:/Users/2021lg003/Documents/LEASYSCAN_ANALYSIS/2-DATA_SMOOTHING/results/PERIODE2/smthTR_FeaturesTimeSeries/D.2023-10-04.csv") 
features_d8<-read.csv("C:/Users/2021lg003/Documents/LEASYSCAN_ANALYSIS/2-DATA_SMOOTHING/results/PERIODE2/smthTR_FeaturesTimeSeries/D.2023-10-05.csv") 
features_d9<-read.csv("C:/Users/2021lg003/Documents/LEASYSCAN_ANALYSIS/2-DATA_SMOOTHING/results/PERIODE2/smthTR_FeaturesTimeSeries/D.2023-10-06.csv") 
features_d10<-read.csv("C:/Users/2021lg003/Documents/LEASYSCAN_ANALYSIS/2-DATA_SMOOTHING/results/PERIODE2/smthTR_FeaturesTimeSeries/D.2023-10-07.csv") 
features_d11<-read.csv("C:/Users/2021lg003/Documents/LEASYSCAN_ANALYSIS/2-DATA_SMOOTHING/results/PERIODE2/smthTR_FeaturesTimeSeries/D.2023-10-08.csv") 
features_d12<-read.csv("C:/Users/2021lg003/Documents/LEASYSCAN_ANALYSIS/2-DATA_SMOOTHING/results/PERIODE2/smthTR_FeaturesTimeSeries/D.2023-10-09.csv") 
features_d13<-read.csv("C:/Users/2021lg003/Documents/LEASYSCAN_ANALYSIS/2-DATA_SMOOTHING/results/PERIODE2/smthTR_FeaturesTimeSeries/D.2023-10-10.csv") 
features_d14<-read.csv("C:/Users/2021lg003/Documents/LEASYSCAN_ANALYSIS/2-DATA_SMOOTHING/results/PERIODE3/smthTR_FeaturesTimeSeries/D.2023-10-11.csv") 
features_d15<-read.csv("C:/Users/2021lg003/Documents/LEASYSCAN_ANALYSIS/2-DATA_SMOOTHING/results/PERIODE3/smthTR_FeaturesTimeSeries/D.2023-10-12.csv") 
features_d16<-read.csv("C:/Users/2021lg003/Documents/LEASYSCAN_ANALYSIS/2-DATA_SMOOTHING/results/PERIODE3/smthTR_FeaturesTimeSeries/D.2023-10-13.csv") 
features_d17<-read.csv("C:/Users/2021lg003/Documents/LEASYSCAN_ANALYSIS/2-DATA_SMOOTHING/results/PERIODE3/smthTR_FeaturesTimeSeries/D.2023-10-14.csv") 
features_d18<-read.csv("C:/Users/2021lg003/Documents/LEASYSCAN_ANALYSIS/2-DATA_SMOOTHING/results/PERIODE3/smthTR_FeaturesTimeSeries/D.2023-10-15.csv") 
features_d19<-read.csv("C:/Users/2021lg003/Documents/LEASYSCAN_ANALYSIS/2-DATA_SMOOTHING/results/PERIODE3/smthTR_FeaturesTimeSeries/D.2023-10-16.csv") 


##rename maxET in MAXET BEFORE BINDING
#Using rename()
features_d1 <- features_d1 %>% 
       rename("MAXET" = "maxET")
features_d2 <- features_d2 %>% 
       rename("MAXET" = "maxET")
# features_d5 <- features_d5 %>% 
#        rename("MAXET" = "maxET")
features_d6 <- features_d6 %>% 
       rename("MAXET" = "maxET")
features_d7 <- features_d7 %>% 
       rename("MAXET" = "maxET")
features_d8 <- features_d8 %>% 
       rename("MAXET" = "maxET")
features_d9 <- features_d9 %>% 
       rename("MAXET" = "maxET")
features_d10 <- features_d10 %>% 
       rename("MAXET" = "maxET")
features_d11 <- features_d11 %>% 
       rename("MAXET" = "maxET")
features_d12 <- features_d12 %>% 
       rename("MAXET" = "maxET")
features_d13 <- features_d13 %>% 
       rename("MAXET" = "maxET")
features_d14 <- features_d14 %>% 
       rename("MAXET" = "maxET")
features_d15<- features_d15 %>% 
       rename("MAXET" = "maxET")
features_d16<- features_d16 %>% 
       rename("MAXET" = "maxET")
features_d17 <- features_d17 %>% 
       rename("MAXET" = "maxET")
features_d18 <- features_d18 %>% 
       rename("MAXET" = "maxET")
features_d19<- features_d19 %>% 
       rename("MAXET" = "maxET")




###ADD DATE ON EACH FEATURE BEFORE BINDING
colnames(features_d1)[8:ncol(features_d1)] <- paste(colnames(features_d1)[8:ncol(features_d1)], '2023-09-28', sep='_')
colnames(features_d2)[8:ncol(features_d2)] <- paste(colnames(features_d2)[8:ncol(features_d2)], '2023-09-29', sep='_')
colnames(features_d5)[8:ncol(features_d5)] <- paste(colnames(features_d5)[8:ncol(features_d5)], '2023-10-02', sep='_')
colnames(features_d6)[8:ncol(features_d6)] <- paste(colnames(features_d6)[8:ncol(features_d6)], '2023-10-03', sep='_')
colnames(features_d7)[8:ncol(features_d7)] <- paste(colnames(features_d7)[8:ncol(features_d7)], '2023-10-04', sep='_')
colnames(features_d8)[8:ncol(features_d8)] <- paste(colnames(features_d8)[8:ncol(features_d8)], '2023-10-05', sep='_')
colnames(features_d9)[8:ncol(features_d9)] <- paste(colnames(features_d9)[8:ncol(features_d9)], '2023-10-06', sep='_')
colnames(features_d10)[8:ncol(features_d10)] <- paste(colnames(features_d10)[8:ncol(features_d10)], '2023-10-07', sep='_')
colnames(features_d11)[8:ncol(features_d11)] <- paste(colnames(features_d11)[8:ncol(features_d11)], '2023-10-08', sep='_')
colnames(features_d12)[8:ncol(features_d12)] <- paste(colnames(features_d12)[8:ncol(features_d12)], '2023-10-09', sep='_')
colnames(features_d13)[8:ncol(features_d13)] <- paste(colnames(features_d13)[8:ncol(features_d13)], '2023-10-10', sep='_')
colnames(features_d14)[8:ncol(features_d14)] <- paste(colnames(features_d14)[8:ncol(features_d14)], '2023-10-11', sep='_')
colnames(features_d15)[8:ncol(features_d15)] <- paste(colnames(features_d15)[8:ncol(features_d15)], '2023-10-12', sep='_')
colnames(features_d16)[8:ncol(features_d16)] <- paste(colnames(features_d16)[8:ncol(features_d16)], '2023-10-13', sep='_')
colnames(features_d17)[8:ncol(features_d17)] <- paste(colnames(features_d17)[8:ncol(features_d17)], '2023-10-14', sep='_')
colnames(features_d18)[8:ncol(features_d18)] <- paste(colnames(features_d18)[8:ncol(features_d18)], '2023-10-15', sep='_')
colnames(features_d19)[8:ncol(features_d19)] <- paste(colnames(features_d19)[8:ncol(features_d19)], '2023-10-16', sep='_')


##differents lengths of features dataframe :

features_d1$Geno_Trt<-paste(features_d1$Genotype, "|", features_d1$Treatment)  



features1=cbind(features_d1,features_d2)
write.table(features1, "Features_PERIODE1.csv", sep = ";")

features2=cbind(features_d5, features_d6, features_d7, features_d8, features_d9, features_d10, features_d11, features_d12, features_d13 )
 write.table(features2, "Features_PERIODE2.csv", sep = ";")
 
features3=cbind(features_d14   , features_d15, features_d16, features_d17, features_d18, features_d19)             
  write.table(features3, "Features_PERIODE3.csv", sep = ";", dec = ".")
          
#################################"""
  
features= read.csv2("FEATURES.csv", sep = ";", dec = ".")
features_info=features[3:9]

char_columns <- sapply(features, is.character)             # Identify character columns
data_chars_as_num <- features                              # Replicate data
data_chars_as_num[ , char_columns] <- as.data.frame(   # Recode characters as numeric
  apply(data_chars_as_num[ , char_columns], 2, as.numeric))
sapply(data_chars_as_num, class)  
features=cbind(features_info, data_chars_as_num[10:ncol(features)])


MAXET=features[,grepl("MAXET",names(features))] ##probleme : on recupere les features maxET mais aussi slope.maxET,  slope.07.maxET, , curvmaxET .... :( Change before maxET to MAXET!

slope.seven.maxET=features[,grepl("slope.07.maxET",names(features))]         ##ok
slope.zero.seven = features[,grepl("slope.00.07",names(features))]             ##ok
slope.nineteen.midnight = features[,grepl("slope.19.23.45",names(features))] ##ok
curvmaxET = features[,grepl("curvmaxET",names(features))]                  ##ok    
total.auc = features[,grepl("total.auc",names(features))]                   ##ok
auc.ten.fifteen = features[,grepl("auc.10.15",names(features))]             ##ok
sd.ten.fifteen = features[,grepl("sd.10.15",names(features))]                  ##ok
auc.prop.ten.fifteen = features[,grepl("auc.prop.10.15",names(features))]    ##ok
auc.seven.nineteen = features[,grepl("auc.07.19",names(features))]           ##ok
sd.seven.nineteen = features[,grepl("sd.07.19",names(features))]              ##ok
auc.prop.seven.nineteen = features[,grepl("auc.prop.07.19",names(features))]   ##ok
auc.night= features[,grepl("auc.night",names(features))]                         ##ok
cos.sim.index= features[,grepl("cos.sim.index",names(features))]                 ##ok
```


```{r 1 MaxET data }
###put feature on same column
MAXET= cbind(MAXET, features$Geno_Trt, features$Genotype, features$Treatment)

MAXET <- melt(MAXET ,  id.vars = "features$Geno_Trt", variable.name = 'rep')
MAXET$value<-as.numeric(MAXET$value)
MAXET= na.omit(MAXET)
colnames(MAXET)=c( "Geno_Trt", "Date", "value")            
MAXET$Date<-gsub("MAXET_","",as.character(MAXET$Date))##remove MAXET before date 
MAXET$Date=lubridate::ymd(MAXET$Date)
 MAXET = separate_wider_delim(MAXET , cols = Geno_Trt, delim = "|", names = c("Geno", "Trt"))
 colnames(MAXET)=c("Genotype", "Treatment","Date", "maxET")
 
 MAXET$Genotype=as.factor(MAXET$Genotype)
 uniq_date=unique(MAXET$Date)
 

 
```


```{r 1 anova maxET data }

 for (d in 1:length(uniq_date)) {
   
   subset_feat= MAXET[MAXET$Date == uniq_date[d],]### pick a day 

   model=lm(maxET~Genotype, data = subset_feat)
  aov<-Anova(model)
  print(aov)
   H2=(aov$`Sum Sq`[1]/(aov$`Sum Sq`[1]+aov$`Sum Sq`[2]))
  print(H2)
   print(d)
   
 }


model=lm(maxET~Genotype, data = MAXET)
aov<-Anova(model)
print(aov)
```


```{r 1 H2 maxET }
H2=(aov$`Sum Sq`[1]/(aov$`Sum Sq`[1]+aov$`Sum Sq`[2]))
H2
# 0.09174557
```


```{r 2 TOTAL AUC  data }
###put feature on same column
total.auc= cbind(total.auc, features$Geno_Trt, features$Genotype, features$Treatment)

total.auc <- melt(total.auc ,  id.vars = "features$Geno_Trt", variable.name = 'rep')
total.auc$value<-as.numeric(total.auc$value)
total.auc= na.omit(total.auc)
colnames(total.auc)=c( "Geno_Trt", "Date", "value")            
total.auc$Date<-gsub("total.auc_","",as.character(total.auc$Date))##remove total.auc before date 
total.auc$Date=lubridate::ymd(total.auc$Date)
 total.auc = separate_wider_delim(total.auc , cols = Geno_Trt, delim = "|", names = c("Geno", "Trt"))
 colnames(total.auc)=c("Genotype", "Treatment","Date", "total.auc")
```


```{r 1 anova total auc data }
model=lm(total.auc~Genotype, data = total.auc)
aov<-Anova(model)
print(aov)

# Response: total.auc
#              Sum Sq    Df F value    Pr(>F)    
# Genotype  1.948e+17   313  2.0955 < 2.2e-16 ***
# Residuals 3.683e+18 12401                    
```

```{r}
H2=(aov$`Sum Sq`[1]/(aov$`Sum Sq`[1]+aov$`Sum Sq`[2]))
H2
# [1] 0.05023427
```



```{r TR   data }

TR = cbind(TR, features$Geno_Trt, features$Genotype, features$Treatment)

TR <- melt(TR ,  id.vars = "features$Geno_Trt", variable.name = 'rep')
TR$value<-as.numeric(TR$value)
TR= na.omit(TR)
colnames(TR)=c( "Geno_Trt", "Date", "value")            
TR$Date<-gsub("TR_","",as.character(TR$Date))
TR$Date=lubridate::ymd(TR$Date)
 TR = separate_wider_delim(TR , cols = Geno_Trt, delim = "|", names = c("Geno", "Trt"))
 colnames(TR)=c("Genotype", "Treatment","Date", "TR")
 
```





```{r 1 anova TR - genotype data }
model=lm(TR~Genotype, data = TR)
aov<-Anova(model)
print(aov)

                  
```

```{r Hé }
H2=(aov$`Sum Sq`[1]/(aov$`Sum Sq`[1]+aov$`Sum Sq`[2]))
H2
```

